Splitting It Up: The \pkg{spduration} Split-Population Duration Regression Package

### 2016-03-08 manuscript2.tex
added cites for smcure

Alltogether the data include 290 coups, which makes for a positive rate of slightly more than 5\% of all observations.

@@

The \pkg{survival} library \citep{thernau2000modeling, thernau2015survival}, arguable the most well-established library for survival and event history analysis in \proglang{R}, allows for the estimation of a wide variet of semi-, non-, and parametric survival models, and provides facilities for handling and descriptive summaries of survival and event history data. It does not include cure rate or split-population mixture models of the type we have implemented though.

@@

Otherwise, the \pkg{smcure} package \citep{cai2012smcureR}, described in \citet{cai2012smcure}, appears to be the only alternative split-population/cure rate duration model implementation in \proglang{R}. It implements semi-parametric proportional hazards and accelerated failure time cure models. The semi-paramatric estimation of the hazard function shape obviates the need to make an assumption about its shape, which can be difficult in practice to empirically evalute, as the \pkg{smcure} discuss. However, it is much more limited in practical application as it only accomodates case-based data, not time-varying covariates.

Furthermore, \proglang{Stata} implements cure models similar to our own, and which can handle time-varying covariates, but this is of course of limited appeal to \proglang{R} users.

>>>>

Altogether the data include 290 coups, which makes for a positive rate of slightly more than 5\% of all observations.

@@

The \pkg{survival} library \citep{thernau2000modeling, thernau2015survival}, arguably the most well-established library for survival and event history analysis in \proglang{R}, allows for the estimation of a wide variety of semi-, non-, and parametric survival models, and provides facilities for handling and descriptive summaries of survival and event history data. It does not include cure rate or split-population mixture models of the type we have implemented though.

@@

Otherwise, the \pkg{smcure} package \citep{cai2012smcureR}, described in \citet{cai2012smcure}, appears to be the only alternative split-population/cure rate duration model implementation in \proglang{R}. It implements semi-parametric proportional hazards and accelerated failure time cure models using methods outlined in \citep{peng2003fitting} and \citep{zhang2007new}. The semi-parametric estimation of the hazard function shape requires no assumption about its shape, which can be difficult in practice to empirically evaluate, as the authors of \pkg{smcure} discuss. However, \pkg{spduration} offers two advantages over \pkg{smcure}. First, it is much less limited in practical application as it allows for the inclusion of time-varying covariates. Second, it features more post-estimation methods, including those discussed above to create and evaluate model predictions.

Finally, \proglang{Stata} implements cure models similar to our own, and which can handle time-varying covariates, but this is of course of limited appeal to \proglang{R} users.
